MRD_final_microbiome

### 2019-08-01 MRD_final_microbiome .Rmd
updated

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '~/globalmicrobiome')
# test
#test
#test
#clash 1
#clash2maryam

```{r functions}

##### FUNCTIONS ########
##### Normalization #######

@@

@@

#setwd("/Users/maryam/Scripts/Microbiome")

@@

#The phyloseq package seeks to address issues with multiple microbiome analysis packages by providing a set of functions that internally manage the organizing, linking, storing, and analyzing of phylogenetic sequencing data. In general, this package is used for UniFrac analyses.
library(phyloseq)
library(microbiome)
library(knitr)
library(ggpubr)
#library(phylogeo)
library(dplyr)
#Graphing package used in phyloseq. To edit the default setting of a plot, you need to use functions in this package.
library(ggplot2)
library(viridis) 
library(cowplot)
# install.packages("devtools")
library(devtools)
# install_github("opisthokonta/tsnemicrobiota")
library(tsnemicrobiota)
# check you've installed this library
library(forcats)

# Data from http://dx.doi.org/10.1038/ncomms7342
#This data set contains genus-level microbiota profiling with HITChip for 1006 western adults with no reported health complications.

@@

## MND,PD,AD Variable
```{r mnd variable}

# Create mnd variable
# grabs the nationality from phyloseq
mbnational <- get_variable(pseq, "nationality" )

# this collapses the two varialbes into a new variable AB in this case. The command for multiple changes is fct_collapse(x, AB = c("A","B"), DE = c("D","E"))
mndmbnational <- fct_collapse(mbnational, LOW = c("Scandinavia","EasternEurope") , MEDIUM = c("SouthEurope","CentralEurope", "UKIE"), HIGH = c("US") )

pdmbnational <- fct_collapse(mbnational, LOW = c("Scandinavia","EasternEurope") , MEDIUM = c("SouthEurope","CentralEurope", "UKIE"), HIGH = c("US") )

admbnational <- fct_collapse(mbnational, LOW = c("Scandinavia","EasternEurope") , MEDIUM = c("SouthEurope","CentralEurope", "UKIE"), HIGH = c("US") )

# reorder
mndmbnational <- factor(mndmbnational, levels = (c("LOW", "MEDIUM", "HIGH")))
pdmbnational <- factor(pdmbnational, levels = (c("LOW", "MEDIUM", "HIGH")))
admbnational <- factor(admbnational, levels = (c("LOW", "MEDIUM", "HIGH")))

# creates a new variable in the phyloseq called mnd
sample_data(pseq)$mnd = mndmbnational
sample_data(pseq)$pd = pdmbnational
sample_data(pseq)$ad = admbnational

# checks that it has worked.
#get_variable(pseq, "mnd")

@@

clostridium <- subset_taxa(pseq, Phylum =="Clostridium cluster I" | Phylum =="Clostridium cluster XI"| Phylum =="Clostridium cluster XVI"  | Phylum =="Clostridium cluster III"  | Phylum =="Clostridium cluster XIII" | Phylum =="Clostridium cluster XVII"| Phylum =="Clostridium cluster IV" | Phylum =="Clostridium cluster XVIa" | Phylum =="Clostridium cluster XVIII" | Phylum =="Clostridium cluster IX" | Phylum =="Clostridium cluster XV" | Phylum =="Uncultured Clostridiales")

pseq.new <- subset_taxa(pseq, Phylum =="Actinobacteria" | Phylum =="Bacilli"| Phylum =="Proteobacteria"  | Phylum =="Verruccomicrobia"  | Phylum =="Bacteroidetes" | Phylum =="Asteroleplasma"| Phylum =="Spirochaetes" | Phylum =="Fusobacteria" | Phylum =="Cyanobacteria" | Phylum =="Uncultured Mollucutes")

@@

## pseq summary
```{r pseq}
pseq
```

```{r List of genera in phylum}
# List of Genera in the Bacteroideted Phylum
taxa <- map_levels(NULL, "Phylum", "Genus", pseq)$Bacteroidetes
print(taxa)
```

```{r MND table summary}
table(sample_data(pseq)$nationality)
table(sample_data(pseq)$mnd)
table(sample_data(pseq)$pd)
table(sample_data(pseq)$ad)
```

## Summary REMOVE?
```{r summary}

# Summarise contents of phyloseq
summarize_phyloseq(pseq)
# Sparsity is how populated is the data with zeros
kable(head(tax_table(pseq)))
tax_table(pseq)

@@

@@

# Prepare data for vizualisation
# Alpha diversity for the entire data set
png("alpha.png")
tab <- diversities(pseq, index = "all")
stargazer(tab, type = "text",summary = NULL)
kable(head(tab))
plot(tab)
dev.off()

@@

pseq.meta$pd <- factor(pseq.meta$pd, levels = c("LOW","MEDIUM","HIGH"),  exclude = "NA")

pseq.meta$ad <- factor(pseq.meta$ad, levels = c("LOW","MEDIUM","HIGH"),  exclude = "NA")

# Create a list of pairwise comaprisons
mnd <- levels(pseq.meta$mnd) # get the variables
pd <- levels(pseq.meta$pd) # get the variables
ad <- levels(pseq.meta$ad) # get the variables

@@

pd.pairs <- combn(seq_along(pd), 2, simplify = FALSE, FUN = function(i)pd[i])
pd.pairs <- list( c("LOW","MEDIUM"), c("HIGH","LOW"), c("MEDIUM","HIGH"))

ad.pairs <- combn(seq_along(ad), 2, simplify = FALSE, FUN = function(i)ad[i])
ad.pairs <- list( c("LOW","MEDIUM"), c("HIGH","LOW"), c("MEDIUM","HIGH"))

# Shannon
# the number in the palettes must equal the number of groups in bmi.pairs
p1 <- ggviolin(pseq.meta, x = "mnd", y = "Shannon", add = "boxplot", fill = "mnd", palette = c("yellow", "orange", "red"))
p1 <- p1 + stat_compare_means(comparisons = mnd.pairs, paired = FALSE) + xlab("MND Incidence") + theme(legend.position="none")
print(p1)
ggsave("mnd_violin__shannon_stats.png")

# Inverse Simpson
p2 <- ggviolin(pseq.meta, x = "mnd", y = "InverseSimpson", add = "boxplot", fill = "mnd", palette = c("yellow", "orange", "red"))
p2 <- p2 + stat_compare_means(comparisons = mnd.pairs) 
print(p2)
ggsave("mnd_violin_simpson.png")

@@

png("Phylum_NMDS.png")
pseq.ord <- ordinate(pseq, "NMDS", "bray")
p1 = plot_ordination(pseq, pseq.ord, type = "taxa", color = "mnd", title = "Taxa by Phylum")
p1 + facet_wrap(~Phylum, 3)
dev.off()

@@

 The low group tends to have smaller values, indicating that the samples are more similar to the group mean, and is less heterogeneous (has smaller spread / is more homogeneous):

@@

p1 = plot_ordination(pseq.rel, pseq.ord, type = "sample", color = "mnd")
plot(p1 + scale_colour_manual(values = c("yellow", "orange", "red", "grey")) + labs(color = "MND Prevalence"))
ggsave("NMDStMND.png")

p1 + scale_colour_manual(values = c("yellow", "orange", "red", "grey")) + facet_wrap(~mnd, nrow = 2) + labs(color = "MND Prevalence")
ggsave("NMDS_facetMND.png")

p2 = plot_ordination(pseq.rel, pseq.ord, type = "sample", color = "pd")
plot(p2 + scale_colour_manual(values = c("yellow", "orange", "red", "grey")) + labs(color = "PD Prevalence"))
ggsave("NMDStpd.png")

p2 + scale_colour_manual(values = c("yellow", "orange", "red", "grey")) + facet_wrap(~pd, nrow = 2) + labs(color = "PD Prevalence")
ggsave("NMDS_facetpd.png")

p3 = plot_ordination(pseq.rel, pseq.ord, type = "sample", color = "ad")
plot(p3 + scale_colour_manual(values = c("yellow", "orange", "red", "grey")) + labs(color = "AD Prevalence"))
ggsave("NMDStad.png")

p3 + scale_colour_manual(values = c("yellow", "orange", "red", "grey")) + facet_wrap(~ad, nrow = 2) + labs(color = "AD Prevalence")
ggsave("NMDS_facetad.png")

@@

p1 + scale_colour_manual(values = c("yellow", "orange", "red", "grey")) + facet_wrap(~mnd, 3) + labs(color = "MND Prevalence")
ggsave("NMDS_facet.png")

p2 = plot_ordination(pseq.rel, pseq.ord, type = "sample", color = "pd")
plot(p2 + scale_colour_manual(values = c( "yellow", "orange", "red", "grey")) + labs(color = "PD Prevalence"))
ggsave("MDSpd.png")

p2 + scale_colour_manual(values = c("yellow", "orange", "red", "grey")) + facet_wrap(~pd, 3) + labs(color = "PD Prevalence")
ggsave("NMDS_facetPD.png")

p3 = plot_ordination(pseq.rel, pseq.ord, type = "sample", color = "ad")
plot(p3 + scale_colour_manual(values = c( "yellow", "orange", "red", "grey")) + labs(color = "AD Prevalence"))
ggsave("MDSad.png")

p3 + scale_colour_manual(values = c("yellow", "orange", "red", "grey")) + facet_wrap(~ad, 3) + labs(color = "AD Prevalence")
ggsave("NMDS_facetAD.png")

@@

t-SNE is a popular new ordination technique
```{r}

tsne_res <- tsne_phyloseq(pseq, distance='bray', perplexity = 5, verbose=1, rng_seed = 3901)
#MND
tsne_p <- plot_tsne_phyloseq(pseq, tsne_res,
  color = 'mnd', title='t-SNE (Bray)')
tsne_p + scale_colour_manual(values = c( "yellow", "orange", "red", "grey")) + labs(color = "MND Prevalence")
ggsave("mnd_t_sne.png")
# PD
tsne_p <- plot_tsne_phyloseq(pseq, tsne_res,
  color = 'pd', title='t-SNE (Bray)')
tsne_p + scale_colour_manual(values = c( "yellow", "orange", "red", "grey")) + labs(color = "PD Prevalence")
ggsave("pd_t_sne.png")

# AD
tsne_p <- plot_tsne_phyloseq(pseq, tsne_res,
  color = 'ad', title='t-SNE (Bray)')
tsne_p + scale_colour_manual(values = c( "yellow", "orange", "red", "grey")) + labs(color = "AD Prevalence")
ggsave("ad_t_sne.png")

#MND
mnd.p.low <- subset_samples(pseq, mnd == "LOW")
mnd.p.low.d <- divergence(mnd.p.low)

mnd.p.med <- subset_samples(pseq, mnd == "MEDIUM")
p.med.d <- divergence(mnd.p.med)

mnd.p.high <- subset_samples(pseq, mnd == "HIGH")
p.high.d <- divergence(mnd.p.high)

mnd.p <- boxplot(list(Low = mnd.p.low.d, Medium = mnd.p.med.d, High = mnd.p.high.d))
plot(mnd.p)

# PD

pd.p.low <- subset_samples(pseq, pd == "LOW")
pd.p.low.d <- divergence(pd.p.low)

pd.p.med <- subset_samples(pseq, pd == "MEDIUM")
p.med.d <- divergence(pd.p.med)

pd.p.high <- subset_samples(pseq, pd == "HIGH")
p.high.d <- divergence(pd.p.high)

pd.p <- boxplot(list(Low = pd.p.low.d, Medium = pd.p.med.d, High = pd.p.high.d))
plot(pd.p)

# AD

ad.p.low <- subset_samples(pseq, ad == "LOW")
ad.p.low.d <- divergence(ad.p.low)

ad.p.med <- subset_samples(pseq, ad == "MEDIUM")
p.med.d <- divergence(ad.p.med)

ad.p.high <- subset_samples(pseq, ad == "HIGH")
p.high.d <- divergence(ad.p.high)

ad.p <- boxplot(list(Low = ad.p.low.d, Medium = ad.p.med.d, High = ad.p.high.d))
plot(ad.p)

@@

# Convert to compositional data
# MND
mnd.pseqLvH <- merge_phyloseq(mnd.p.low,mnd.p.high)
mnd.pseqLvM <- merge_phyloseq(mnd.p.low,mnd.p.med)

mnd.pseq.rel <- microbiome::transform(mnd.pseqLvH, "compositional")
mnd.otu <- abundances(mnd.pseq.rel)
mnd.meta <- meta(mnd.pseq.rel)

# PD

pd.pseqLvH <- merge_phyloseq(pd.p.low,pd.p.high)
pd.pseqLvM <- merge_phyloseq(pd.p.low,pd.p.med)

pd.pseq.rel <- microbiome::transform(pd.pseqLvH, "compositional")
pd.otu <- abundances(pd.pseq.rel)
pd.meta <- meta(pd.pseq.rel)

# AD

ad.pseqLvH <- merge_phyloseq(ad.p.low,ad.p.high)
ad.pseqLvM <- merge_phyloseq(ad.p.low,ad.p.med)

ad.pseq.rel <- microbiome::transform(ad.pseqLvH, "compositional")
ad.otu <- abundances(ad.pseq.rel)
ad.meta <- meta(ad.pseq.rel)

@@

# samples x species as input
library(vegan)

# MND
mnd.permanova <- adonis(t(mnd.otu) ~ mnd,
               data = meta, permutations=999, method = "bray")
# P-value
print(as.data.frame(mnd.permanova$aov.tab)["mnd", "Pr(>F)"])
summary(mnd.permanova$aov.tab)

# PD
pd.permanova <- adonis(t(pd.otu) ~ pd,
               data = meta, permutations=999, method = "bray")
# P-value
print(as.data.frame(pd.permanova$aov.tab)["pd", "Pr(>F)"])
summary(pd.permanova$aov.tab)

# AD
ad.permanova <- adonis(t(ad.otu) ~ ad,
               data = meta, permutations=999, method = "bray")
# P-value
print(as.data.frame(ad.permanova$aov.tab)["ad", "Pr(>F)"])
summary(ad.permanova$aov.tab)

@@

Check that variance homogeneity assumptions hold (to ensure the reliability of the results):Check that variance homogeneity assumptions hold (to ensure the reliability of the results):
```{r ANOVA}
#MND
mnd.dist <- vegdist(t(mnd.otu))
mnd.anov <- anova(betadisper(mnd.dist, meta$mnd))
summary(mnd.anov)
tidy(mnd.anov)

#PD
pd.dist <- vegdist(t(pd.otu))
pd.anov <- anova(betadisper(pd.dist, meta$pd))
summary(pd.anov)
tidy(pd.anov)

#AD
ad.dist <- vegdist(t(ad.otu))
ad.anov <- anova(betadisper(ad.dist, meta$ad))
summary(ad.anov)
tidy(ad.anov)

@@

# MDN
mnd.coef <- coefficients(mnd.permanova)["mnd1",]
mnd.top.coef <- coef[rev(order(abs(mnd.coef)))[1:20]]
par(mar = c(3, 14, 2, 1))
barplot(sort(mnd.top.coef), col = rainbow(20), horiz = T, las = 1, main = "Top Genus") 
ggsave("mndLvH.png")

# PD
pd.coef <- coefficients(pd.permanova)["pd1",]
pd.top.coef <- coef[rev(order(abs(pd.coef)))[1:20]]
par(mar = c(3, 14, 2, 1))
barplot(sort(pd.top.coef), col = rainbow(20), horiz = T, las = 1, main = "Top Genus") 
ggsave("pdLvH.png")

# PD
ad.coef <- coefficients(ad.permanova)["ad1",]
ad.top.coef <- coef[rev(order(abs(ad.coef)))[1:20]]
par(mar = c(3, 14, 2, 1))
barplot(sort(ad.top.coef), col = rainbow(20), horiz = T, las = 1, main = "Top Genus") 
ggsave("adLvH.png")

@@

#MND
mnd.pseqLvM <- merge_phyloseq(mnd.p.low,mnd.p.med)
mnd.pseq.rel <- microbiome::transform(mnd.pseqLvM, "compositional")
mnd.otu <- abundances(mnd.pseq.rel)
mnd.meta <- meta(mnd.pseq.rel)

#PD
pd.pseqLvM <- merge_phyloseq(pd.p.low,pd.p.med)
pd.pseq.rel <- microbiome::transform(pd.pseqLvM, "compositional")
pd.otu <- abundances(pd.pseq.rel)
pd.meta <- meta(pd.pseq.rel)

#AD
ad.pseqLvM <- merge_phyloseq(ad.p.low,ad.p.med)
ad.pseq.rel <- microbiome::transform(ad.pseqLvM, "compositional")
ad.otu <- abundances(ad.pseq.rel)
ad.meta <- meta(ad.pseq.rel)

# MND
library(vegan)
mnd.permanova <- adonis(t(mnd.otu) ~ mnd,
               data = meta, permutations=999, method = "bray")
# P-value
print(as.data.frame(mnd.permanova$aov.tab)["mnd", "Pr(>F)"])
mnd.dist <- vegdist(t(mnd.otu))
mnd.anov <- anova(betadisper(mnd.dist, mnd.meta$mnd))
summary(mnd.anov)
tidy(mnd.anov)
mnd.coef <- coefficients(mnd.permanova)["mnd1",]
mnd.top.coef <- coef[rev(order(abs(mnd.coef)))[1:20]]
par(mar = c(3, 14, 2, 1))
barplot(sort(mnd.top.coef), col = rainbow(20), horiz = T, las = 1, font.axis = 3, main = "Top Genus")
ggsave("mndLvM.png", width = 20, height = 20, units = "cm")

# PD
library(vegan)
pd.permanova <- adonis(t(pd.otu) ~ pd,
               data = meta, permutations=999, method = "bray")
# P-value
print(as.data.frame(pd.permanova$aov.tab)["pd", "Pr(>F)"])
pd.dist <- vegdist(t(pd.otu))
pd.anov <- anova(betadisper(pd.dist, pd.meta$pd))
summary(pd.anov)
tidy(pd.anov)
pd.coef <- coefficients(pd.permanova)["pd1",]
pd.top.coef <- coef[rev(order(abs(pd.coef)))[1:20]]
par(mar = c(3, 14, 2, 1))
barplot(sort(pd.top.coef), col = rainbow(20), horiz = T, las = 1, font.axis = 3, main = "Top Genus")
ggsave("pdLvM.png", width = 20, height = 20, units = "cm")

# AD
library(vegan)
ad.permanova <- adonis(t(ad.otu) ~ ad,
               data = meta, permutations=999, method = "bray")
# P-value
print(as.data.frame(ad.permanova$aov.tab)["ad", "Pr(>F)"])
ad.dist <- vegdist(t(ad.otu))
ad.anov <- anova(betadisper(ad.dist, ad.meta$ad))
summary(ad.anov)
tidy(ad.anov)
ad.coef <- coefficients(ad.permanova)["ad1",]
ad.top.coef <- coef[rev(order(abs(ad.coef)))[1:20]]
par(mar = c(3, 14, 2, 1))
barplot(sort(ad.top.coef), col = rainbow(20), horiz = T, las = 1, font.axis = 3, main = "Top Genus")
ggsave("adLvM.png", width = 20, height = 20, units = "cm")

@@

png("Phylum_abundance.png")
p<- plot_bar(pseq,x="Phylum",fill = "Phylum") +geom_bar(stat="identity")

phylumGlommed = tax_glom(pseq, "Phylum")
plot_bar(phylumGlommed, x="Phylum", fill ="Phylum") +geom_bar(stat="identity")

pd <- p$data %>% 
    as_tibble %>%
    mutate(Genus = as.character(Genus))
genus_abun <- pd %>%
    group_by(Genus) %>%
    summarize(Abundance = sum(Abundance)) %>%
    arrange(Abundance)
genus_levels <- genus_abun$Genus
pd0 <- pd %>%
    mutate(Genus = factor(Genus, genus_levels))
ggplot(pd0, aes(x = Sample, y = Abundance, fill = Genus)) + 
    geom_bar(stat = "identity")
dev.off()

@@

theme(axis.text.x = element_text(angle = 90)
+ guide_italics
  ggtitle("Relative abundance")  + 
```{r Relative abundance}

@@

plot.comp.rel <- plot_composition(pseq.ph.rel, x.label = "sample") +
  theme(legend.position = "bottom") + theme_bw() + 
  theme(axis.text.x=element_blank()) + 
  theme(legend.title = element_text(size=18)) + scale_fill_brewer(palette = "Paired")
print(plot.comp.rel)
ggsave("phylum_relative_abundance11.png")

@@

plot.comp.rel <- plot_composition(clos.ph.rel, x.label = "sample") + 
  theme(legend.position = "bottom", ) + theme_bw() + 
  theme(axis.text.x=element_blank()) + 
  theme(legend.title = element_text(size=18))
plot.comp.rel + scale_fill_brewer(palette = "Paired")

plot.comp.rel <- plot_composition(pseq.ph.rel, x.label = "sample") +
 theme(legend.position = "bottom") + theme_bw() +
 theme(axis.text.x=element_blank()) +
 theme(legend.title = element_text(size=18),legend.text = element_text( face="italic") ) + scale_fill_brewer(palette = "Paired") + guides(fill=guide_legend("my awesome title"))

@@

plot.comp.rel + scale_fill_brewer(palette = "Paired") + theme(legend.position="none")
ggsave("phylum_relative_abundance_clos2.png")

@@

pseq.ph <- aggregate_taxa(pseq, "Phylum")
pseq.ph.rel <- microbiome::transform(pseq.ph, "compositional")
#MND
plot.comp.rel <- plot_composition(pseq.ph.rel, average_by = "mnd", x.label = "MND Incidence", plot.type = "barplot")
plot.comp.rel + scale_fill_viridis(discrete = TRUE, option = "plasma") 
ggsave("mndphylum_abundance_bar.png")
#PD
plot.comp.rel <- plot_composition(pseq.ph.rel, average_by = "pd", x.label = "PD Incidence", plot.type = "barplot")
plot.comp.rel + scale_fill_viridis(discrete = TRUE, option = "plasma") 
ggsave("pdphylum_abundance_bar.png")
#AD
plot.comp.rel <- plot_composition(pseq.ph.rel, average_by = "ad", x.label = "AD Incidence", plot.type = "barplot")
plot.comp.rel + scale_fill_viridis(discrete = TRUE, option = "plasma") 
ggsave("adphylum_abundance_bar.png")
```

@@

p.new.rel <- microbiome::transform(p.new, "compositional")
#MND
plot.p.new.rel <- plot_composition(p.new.rel, average_by = "mnd", x.label = "sample", plot.type = "barplot")
plot.p.new.rel + scale_fill_viridis(discrete = TRUE, option = "plasma") 
ggsave("mndpnew_abundance_bar.png")
#PD
plot.p.new.rel <- plot_composition(p.new.rel, average_by = "pd", x.label = "sample", plot.type = "barplot")
plot.p.new.rel + scale_fill_viridis(discrete = TRUE, option = "plasma") 
ggsave("pdpnew_abundance_bar.png")
#AD
plot.p.new.rel <- plot_composition(p.new.rel, average_by = "ad", x.label = "sample", plot.type = "barplot")
plot.p.new.rel + scale_fill_viridis(discrete = TRUE, option = "plasma") 
ggsave("adpnew_abundance_bar.png")

@@

# Plot - Phylum PD
p.ra.phylum <- ggplot(pseq_phylum, aes(x = sample, y = Abundance, fill = Phylum)) + 
  geom_bar(stat = "identity", width = 1) +
  facet_wrap(~pd, scales = "free_x", nrow = 4, ncol = 7) +
  theme(axis.text.x = element_blank()) +
  theme(axis.title.x = element_blank()) +
  labs(title = "Abundant Phylum (> 1%)")

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##### FUNCTIONS ########

##### Normalization #######

@@

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

Atlas1006
https://rdrr.io/github/microbiome/microbiome/man/atlas1006.html
Using dataset: atlas1006: HITChip Atlas with 1006 Western adults
This data set contains genus-level microbiota profiling with HITChip for 1006 western adults with no reported health complications. 
Format: phyloseq-class format
Usage: data(atlas1006)

@@

setwd("/Users/maryam/Scripts/Microbiome")

@@

#The phyloseq package seeks to address issues with multiple microbiome analysis packages by providing a set of functions that internally manage the organizing, linking, storing, and analyzing of phylogenetic sequencing data. In general, this package is used for UniFrac analyses.
library(phyloseq)
library(microbiome)
library(knitr)
library(ggpubr)

  #library(phylogeo)
library(dplyr)

#Graphing package used in phyloseq. To edit the default setting of a plot, you need to use functions in this package.
library(ggplot2)

library(viridis) 
library(cowplot)

# install.packages("devtools")
library(devtools)
# install_github("opisthokonta/tsnemicrobiota")
library(tsnemicrobiota)

# Data from http://dx.doi.org/10.1038/ncomms7342
data(atlas1006)

@@

## MND Variable
```{r mnd variable}

# Create mnd variable
# grabs the nationality from phyloseq
test <- get_variable(pseq, "nationality" )

# check you've installed this library
library(forcats)

# this collapses the two varialbes into a new variable AB in this case. The command for multiple changes is fct_collapse(x, AB = c("A","B"), DE = c("D","E"))
test <- fct_collapse(test, LOW = c("Scandinavia","EasternEurope") , MEDIUM = c("SouthEurope","CentralEurope", "UKIE"), HIGH = c("US") )

# reorder
test <- factor(test, levels = (c("LOW", "MEDIUM", "HIGH")))
levels(test)

# creates a new variable in the phyloseq called mnd
sample_data(pseq)$mnd = test

# checks that it has worked.
get_variable(pseq, "mnd")

@@

clostridium <- subset_taxa(pseq, Phylum =="Clostridium cluster I" | Phylum =="Clostridium cluster XI"| Phylum =="Clostridium cluster XVI"  | Phylum =="Clostridium cluster III"  | Phylum =="Clostridium cluster XIII" | Phylum =="Clostridium cluster XVII"| Phylum =="Clostridium cluster IV" | Phylum =="Clostridium cluster XIVa" | Phylum =="Clostridium cluster XVIII" | Phylum =="Clostridium cluster IX" | Phylum =="Clostridium cluster XV" | Phylum =="Uncultured Clostridiales")

pseq.new <- subset_taxa(pseq, Phylum =="Actinobacteria" | Phylum =="Bacilli"| Phylum =="Proteobacteria"  | Phylum =="Verrucomicrobia"  | Phylum =="Bacteroidetes" | Phylum =="Asteroleplasma"| Phylum =="Spirochaetes" | Phylum =="Fusobacteria" | Phylum =="Cyanobacteria" | Phylum =="Uncultured Mollicutes")

@@

## pseq summary
```{r pseq}

pseq

#### List of genera in phylums
```{r List of genera in phylum}

# List of Genera in the Bacteroideted Phylum
taxa <- map_levels(NULL, "Phylum", "Genus", pseq)$Bacteroidetes
print(taxa)

map_levels(NULL, "Phylum", "Genus", pseq)

```{r MND table summary}

table(sample_data(pseq)$mnd)

table(sample_data(pseq)$nationality)

## Summary
```{r summary}

# Summarise contents of phyloseq
summarize_phyloseq(pseq)
# Sparsity is how populated is the data with zeros

kable(head(tax_table(pseq)))
tax_table(pseq)

@@

```{r average}

library(purrr)
library(pastecs)
pseq.meta %>% split(.$mnd) %>% map(stat.desc)

@@

# Prepare data for vizualisation

# Alpha diversity again
png("alpha.png")
tab <- diversities(pseq, index = "all")
stargazer(tab, type = "text",summary = NULL)
kable(head(tab))
plot(tab)
dev.off()

@@

# Create a list of pairwise comaprisons
mnd <- levels(pseq.meta$mnd) # get the variables

@@

# Shannon
# the number in the palettes must equal the number of groups in bmi.pairs
p1 <- ggviolin(data=subset(pseq.meta, !is.na(mnd)), x = "mnd", y = "Shannon", add = "boxplot", fill = "mnd", na.rm = TRUE, palette = c("yellow", "orange", "red"))
p1 <- p1 + stat_compare_means(comparisons = mnd.pairs) + xlab("MND Prevalence") + theme(legend.position="none")
print(p1)
ggsave("mnd_violin__shannon_stats.png")

# Inverse Simpson
p2 <- ggviolin(data=subset(pseq.meta, !is.na(mnd)), x = "mnd", y = "InverseSimpson", add = "boxplot", fill = "mnd", palette = c("yellow", "orange", "red"))
p2 <- p2 + stat_compare_means(comparisons = mnd.pairs) 
print(p2)
ggsave("mnd_violin_simpson.png")

@@

png("Phylum_NMDS.png")
pseq.ord <- ordinate(pseq, "NMDS", "bray")

p1 = plot_ordination(pseq, pseq.ord, type = "taxa", color = "mnd", title = "Taxa by Phylum")
p1 + facet_wrap(~Phylum, 3)
dev.off()

@@

p.low <- subset_samples(pseq, mnd == "LOW")
p.low.d <- divergence(p.low)

p.med <- subset_samples(pseq, mnd == "MEDIUM")
p.med.d <- divergence(p.med)

p.high <- subset_samples(pseq, mnd == "HIGH")
p.high.d <- divergence(p.high)

p <- boxplot(list(Low = p.low.d, Medium = p.med.d, High = p.high.d))
plot(p)

```
 The low group tends to have smaller values, indicating that the samples are more similar to the group mean, and is less heterogeneous (has smaller spread / is more homogeneous):

@@

p1 = plot_ordination(pseq.rel, pseq.ord, type = "sample", color = "mnd")
plot(p1 + scale_colour_manual(values = c("yellow", "orange", "red", "grey")) + labs(color = "MND Prevalence"))
ggsave("NMDSt.png")

p1 + scale_colour_manual(values = c("yellow", "orange", "red", "grey")) + facet_wrap(~mnd, nrow = 2) + labs(color = "MND Prevalence")
ggsave("NMDS_facet.png")

@@

p1 + scale_colour_manual(values = c("yellow", "orange", "red", "grey")) + facet_wrap(~mnd, 3) + labs(color = "MND Prevalence")
ggsave("MDS_facet.png")

@@

t-SNE is a popular new ordination technique
, title='t-SNE (Bray)'
```{r t-SNE}

tsne_res <- tsne_phyloseq(pseq, distance='bray', perplexity = 5, verbose=1, rng_seed = 3901)
tsne_p <- plot_tsne_phyloseq(pseq, tsne_res,
  color = 'mnd')
tsne_p + scale_colour_manual(values = c( "yellow", "orange", "red", "grey")) + labs(color = "MND Prevalence")
ggsave("t_sne.png")

@@

# Convert to compositional data
pseqLvH <- merge_phyloseq(p.low,p.high)

pseq.rel <- microbiome::transform(pseqLvH, "compositional")
otu <- abundances(pseq.rel)
meta <- meta(pseq.rel)

@@

# samples x species as input
library(vegan)
permanova <- adonis(t(otu) ~ mnd,
               data = meta, permutations=999, method = "bray")

# P-value
print(as.data.frame(permanova$aov.tab)["mnd", "Pr(>F)"])

summary(permanova$aov.tab)

@@

Homogeneity of variance is an assumption underlying both t tests and F tests (analyses of variance, ANOVAs) in which the population variances (i.e., the distribution, or “spread,” of scores around the mean) of two or more samples are considered equal.
Check that variance homogeneity assumptions hold (to ensure the reliability of the results):
```{r ANOVA}

dist <- vegdist(t(otu))
anov <- anova(betadisper(dist, meta$mnd))
summary(anov)
tidy(anov)
summary(anov$aov.tab)

@@

coef <- coefficients(permanova)["mnd1",]
top.coef <- coef[rev(order(abs(coef)))[1:20]]
par(mar = c(3, 14, 2, 1))
barplot(sort(top.coef), col = rainbow(20), horiz = T, las = 1, font.axis=3, main = "Top Genus")
ggsave("LvH.png", width = 20, height = 20, units = "cm", dpi=700)

@@

pseqLvM <- merge_phyloseq(p.low,p.med)

pseq.rel <- microbiome::transform(pseqLvM, "compositional")
otu <- abundances(pseq.rel)
meta <- meta(pseq.rel)

library(vegan)
permanova <- adonis(t(otu) ~ mnd,
               data = meta, permutations=999, method = "bray")

# P-value
print(as.data.frame(permanova$aov.tab)["mnd", "Pr(>F)"])
summary(permanova$aov.tab)

dist <- vegdist(t(otu))
anov <- anova(betadisper(dist, meta$mnd))
summary(anov)
tidy(anov)

coef <- coefficients(permanova)["mnd1",]
top.coef <- coef[rev(order(abs(coef)))[1:20]]
par(mar = c(3, 14, 2, 1))

png(filename="LvvM.png")
barplot(sort(top.coef), col = rainbow(20), horiz = T, las = 1, font.axis=3, main = "Top Genus")
dev.off()

@@

plot_bar(pseq,x="Phylum",fill = "Phylum") +geom_bar(stat="identity") + theme(legend.position="none") + theme(axis.text.x = element_text(face="italic"))
ggsave("Phylum_abundance.png", height=5, width=5, units='in', dpi=700)

@@

```{r Relative abundance}

@@

plot.comp.rel <- plot_composition(pseq.ph.rel, x.label = "sample") +
theme(legend.position = "bottom") + theme_bw() +
theme(axis.text.x=element_blank()) +
c + scale_fill_brewer(palette = "Paired") + guides(fill=guide_legend("Phylum"))

print(plot.comp.rel)

ggsave("phylum_relative_abundance11.png")

@@

plot.comp.rel <- plot_composition(clos.ph.rel, x.label = "sample") + 
  theme(legend.position = "bottom", ) + theme_bw() + 
  theme(axis.text.x=element_blank()) + 
  theme(legend.title = element_text(size=18), legend.text = element_text( face="italic")) + scale_fill_brewer(palette = "Paired") + guides(fill=guide_legend(title="Phylum"))

print(plot.comp.rel)

@@

plot.comp.rel + scale_fill_brewer(palette = "Paired") + theme(legend.position="none") + guide_italics
ggsave("phylum_relative_abundance_clos2.png")

@@

pseq.ph <- aggregate_taxa(pseq, "Phylum")

plot.comp.rel <- plot_composition(pseq.ph.rel, average_by = "mnd", x.label = "MND Incidence", plot.type = "barplot") + guides(fill=guide_legend(title="Phylum"))

plot.comp.rel + scale_fill_viridis(discrete = TRUE, option = "plasma") 
ggsave("phylum_abundance_bar.png")

top5P.names = sort(tapply(taxa_sums(pseq), tax_table(pseq)[, "Genus"], sum), TRUE)[1:5]

top5P = subset_taxa(pseq, Genus %in% names(top5P.names))

pseq.ph.rel <- microbiome::transform(top5P, "compositional")

plot.comp.rel <- plot_composition(pseq.ph.rel, average_by = "mnd", x.label = "MND Incidence", plot.type = "barplot") + guides(fill=guide_legend(title="Phylum"))

plot.comp.rel + scale_fill_viridis(discrete = TRUE, option = "plasma")

top10P.names = sort(tapply(taxa_sums(pseq), tax_table(pseq)[, "Genus"], sum), TRUE)[1:10]

top10P = subset_taxa(pseq, Genus %in% names(top10P.names))

pseq.ph.rel <- microbiome::transform(top10P, "compositional")

plot.comp.rel <- plot_composition(pseq.ph.rel, average_by = "mnd", x.label = "MND Incidence", plot.type = "barplot") + guides(fill=guide_legend(title="Phylum"))

plot.comp.rel + scale_fill_viridis(discrete = TRUE, option = "plasma")

@@

p.new.rel <- microbiome::transform(p.new, "compositional")

plot.p.new.rel <- plot_composition(p.new.rel, average_by = "mnd", x.label = " ", plot.type = "barplot") + guides(fill=guide_legend(title="Phylum")) + theme(legend.text = element_text( face="italic"))

plot.p.new.rel + scale_fill_brewer(palette = "Paired") 
ggsave("pnew_abundance_bar.png")

@@
